---
title: "PA1"
output: html_document
---
##Loading and preprocessing the data
```{r Loading and preprocessing the data}
setwd("C:/Users/marti/OneDrive/03_Data/01_Courses/Data Science Specialization/Reproducible research/Week 2/repdata%2Fdata%2Factivity")
amd <- read.csv("activity.csv")
amd$date <- as.Date(amd$date, "%Y-%m-%d")
```

##What is mean total number of steps taken per day?
```{r total number of steps per day}
dTotal <- tapply(amd$steps, amd$date, sum, na.rm = TRUE)
par(mfrow = c(1,1), mar = c(3,3,3,3))
hist(dTotal)
dMean <- mean(dTotal)
dMedian <- median(dTotal)
```

Mean of total number of steps per day is `r dMean` and meadeian of total number of steps per day is `r dMedian`.

##What is the average daily activity pattern?
```{r daily activity pattern}
iMean <- tapply(amd$steps, amd$interval, mean, na.rm = TRUE)
plot(names(iMean), iMean, type = "l")
maxInterval <- names(iMean)[which.max(iMean)]
```

The interval with maximum average steps across all the days is `r maxInterval`

##Imputing missing values
```{r}
naCount <- sum(as.vector(apply(amd, 1, function(x) length(which(is.na(x))) != 0)))
amdExtended <- amd 
for (i in seq_len(nrow(amd))) {
        if(is.na(amd$steps[i])){
                
                amdExtended$steps[i] <- mean(subset(amd, amd$interval == amd$interval[i])$steps, na.rm = TRUE)
        }
}

dTotalExtended <- tapply(amd$steps, amd$date, sum, na.rm = TRUE)
hist(dTotalExtended)
dMeanExtended <- mean(dTotalExtended)
dMedianExtended <- median(dTotalExtended)
dMeanDiff <- dMean - dMeanExtended
dMedianDiff <- dMedian - dMedianExtended
```

Total number of observations with missing values is `r naCount`.
Missing values were inserted based on the mean of given interval across all days.
the new mean of total number of steps per day is `r dMean` and meadeian of total number of steps per day is `r dMedian`.
The replacement of missing values changed the total mean by `r dMeanDiff` and the tolal median by `r dMedianDiff`.

##Are there differences in activity patterns between weekdays and weekends?
```{r}
wd <- weekdays(amd$date) %in% c("Sunday", "Saturday")
amdExtended$weekday <- wdPattern <- factor(wd, labels = c("week", "weekend"))
amdExtendedSplit <- split(amdExtended, amdExtended$weekday)
res <- lapply(amdExtendedSplit, function(x) tapply(x[,1], x[,3], mean))
par(mfrow = c(1,2), mar = c(4,4,4,4))
plot(names(res[[1]]), res[[1]], main = "Week", xlab = "interval", ylab = "Number of steps", type = "l")
plot(names(res[[2]]), res[[2]], main = "Weekend", xlab = "interval", ylab = "Number of steps", type = "l")
```

